\chapter*{Acknowledgements}
\pagenumbering{roman}
\setcounter{page}{1}
\addcontentsline{toc}{chapter}{Acknowledgements}

Life as a PhD student has various advantages.
For example, as part of my job, I got to visit Portugal, England, Belgium, Germany, South-Korea, and France, and if my planning had been a little better, I could have visited the Czech Republic as well.
For this and other reasons, I would like to thank my promotor Mark van den Brand for hiring me and giving me the opportunity to experience all the benefits of a career as a PhD student.
Additionally, I thank Mark for his guidance and support over the years.

Around the start of the final year of my project, Suzana Andova agreed to be my copromotor.
From that moment on, the rate of my scientific output in terms of papers increased significantly, for which I would like to thank Suzana.
Without her confidence and experience as guidance, finishing this thesis would have taken much more time.

Besides Mark and Suzana, Marcel van Amstel and Anton Wijs were co-authors of some of the papers that form the basis of this thesis.
I enjoyed the discussions we had concerning the research we performed, and I learned a lot from both cooperations.

Furthermore, I would like to thank the reading committee, consisting of Twan Basten from the Eindhoven University of Technology, Peter Mosses from Swansea University, and Arend Rensink from the University of Twente, for reviewing this thesis.

During most of my employment at the Software Engineering and Technology group, I shared an office with Marcel van Amstel, Jeroen Arnoldus, and Zvezdan Proti{\'c}.
Later on, we relocated and were joined by Yanja Dajsuren, Arjan van der Meer, Ulyana Tikhonova, and Bogdan Vasilescu.
I enjoyed the time we spent together in our office, during summer schools, at workshops, and at the Efteling, and I would like to thank all of them for that.
Furthermore, I thank all my other colleagues and former colleagues at the Software Engineering and Technology group for providing a pleasant working environment, and Tom Verhoeff in particular for noticing an oversight with significant consequences in one of the model transformations developed by Marcel and me.

It took some time to finish this thesis after my PhD project ended, and I thank Harold Weffers for allowing me to work on it during my employment at LaQuSo.

Finally, I would like to thank my family, friends, volleyball team mates, and my girlfriend Sandra for their support and the much appreciated distractions they provided.

\bigskip

\noindent
\emph{Luc Engelen}\\
Eindhoven, August 2012 